
We consider four types of random network models and analyze their
connectivity structure. The \textit{anisotropic} and \textit{tuned
anisotropic} network models implement the anisotropy in the
. The \textit{rewired} version of those networks serve as a baseline
. The \textit{distance-dependent} network allows for further
distinction between effects.

\subsection*{Anisotropic networks}

The anisotropic network model is the simplest form of a random network
featuring anisotropy in connectivity. In the model the $N$ nodes of
the network are first distributed uniformly at random on a square of
side length $E$. Each node $v$ is randomly assigned an angle
$\alpha \in [0,2\pi)$. Directed connections are then established from
$v$ to other nodes with distance less than $\frac{w}{2}$ to the
projection ray originating from $v$ with angle $\alpha$
(Fig.~\ref{fig:aniso_net}, \ref{fig:4_net_models}A). We chose
$N=1000$, leaving the ratio $\frac{w}{E}$ as a free parameter of the
model determining the connection density of the network. We selected
$E=\SI{296}{\micro\meter}$ and determined $w=\SI{74.6}%
{\micro\meter}$ to obtain a connection density of $p=11.6$ as in
\cite{Song2005}. For a formal mathematical definition of the
anisotropic network model see supplementary material
(\ref{SI-sec:aniso_model}).


\input{figures/figure_1-2.tex}



\bigskip

\subsection*{Rewired networks}
%
In order to identify connectivity features that are a direct result of
a network's anisotropy in spatial connectivity, we introduce rewired
versions of anisotropic networks. In the rewiring process existing
edges are randomly assigned new targets so that anisotropy is
eliminated while other important connectivity statistics, such as
distance-dependent connectivity and in- and out-degree distributions,
are preserved. For example, Fig.~\ref{fig:4_net_models}A shows the
targets of a node in the original anisotropic network which are to be
rewired. The same node is then shown in Fig.~\ref{fig:4_net_models}B
with its new targets in the rewired network.
%
% We simply call the modified networks rewired networks (Fig.~\ref{fig:4_net_models}B).

The rewiring algorithm has two parameters -- the rewiring fraction
$\eta$ of edges to be rewired and the rewiring margin $\varepsilon$ that
determines how many rewiring targets are on average available. In the
process any given edge of the graph gets rewired with probability
$\eta$. If an edge $e$ with source node $s(e)$ and target node $t(e)$
is to be rewired, at first the Euclidean distance $x$ between $s(e)$
and $t(e)$ is recorded. Then the pool of potential new targets for $e$
are all nodes that have a distance to $s(e)$ larger than
$x-\frac{\varepsilon}{2}$ but smaller than $x+\frac{\varepsilon}{2}$
(Fig.~\ref{fig:4_net_models}E). From the pool of potential new targets
are chosen at random until a target is found that does not already
receive input from $s(e)$, either through a previously rewired edge or
an edge that will not be rewired. The original edge $e$ is then
replaced by a new edge $e'$ from $s(e)$ to the new target
(Fig.~\ref{fig:4_net_models}F). If no such target is available the
edge is discarded and will not appear in the rewired network.

We chose a rewiring margin $\varepsilon / E = 0.05$ to provide a large
enough average number of rewiring targets while closely maintaining
the distance-depedent connecitivty in the network (see
Section~\ref{SI-sec:rewiring_margin}). A rewired network, if not
otherwise specified, refers to a fully rewired version ($\eta=1$) of
the current network.

% to achieve an effective rewiring while closely maintaining the distance-dependent connection probability distribution of the original networks (see SI ??).



% a modified
% network that lacks anisotropy but is otherwise equivalent to the
% original graph is required. This is achieved for a rewiring algorithm


% we introduced a rewiring algorithm that eliminates anisotropy but preserves other important 


% ia network without anisotropy as close as possible to. For this we introduce a rewiring algorithm operating
% on the anisotropic networks.

% Thus, to
% discern features that are truly 

% it is necessary to have a reference network that doesn't
% feature anisotropy, but is otherwise equivalent to the anisotropic
% model.

% For this we introduced a rewiring algorithm on networks that
% preserves first-order connectivity statistics %??  as well as the
% distance-dependent connection probability profile. In this process
% each connection with a spatial distance $x$ between source and target,
% is rewired to a new random target that differs not more than
% $\varepsilon$ from $x$ in spatial distance to the source vertex
% (Fig.~\ref{fig_model} E-F). The resulting network has greatly reduced
% anisotropy in connectivity, resembling that of a distance-dependent
% network (supporting information).

% Mention the outcome of the rewiring. Some very edges are lost in the process! %??

% By rewiring only a section of all connections, a \textit{partial
%   rewiring} produces a state that retains some anisotropy...



% The relative width of the rewiring parameter was chosen as $\varepsilon / E = 0.05$ (??) to achieve an effective rewiring while closely maintaining the distance-dependent connection probability distribution of the original networks (see SI ??).



\input{figures/figure_2.tex}

\subsection*{Distance-dependent networks}

A further reference network model are distance-dependent networks, in which each possible connection is realized independently with a probability $p(x)$, that depends only on the Euclidean distance $x$ between the source and target node (Fig.~\ref{fig:4_net_models}C). We refer to $p(x)$ as the distance-dependent connectivity profile.


\subsection*{Tuned anisotropic networks}




