\clearpage
\pagebreak
\newpage
\subsection*{Three neuron motifs}

Multiple paired whole-cell recordings in slices of rat visual and
somatosensory cortex revealed nonrandom structures in layer 5
pyramidal cells beyond pair connections -- the connection motifs
occurring in groups of three neurons deviate significantly of what one
would predict from the statistics of connections in neuron
pairs~\cite{Song2005, Perin2011}. Song et al.~for example found the
probability for a neuron pair $(u,v)$ to have a connection
$u \rightarrow v$ to be $p_{u \rightarrow v}= 0.0615$ and the
probability for a bidirectional connection was
$p_{u \leftrightarrow v} = 0.0542$. Form this they computed the
probability for a group of three randomly selected neurons to show,
for example, the connection motif 14 as labelled in
\figref{fig:3_motifs}A as
$p_{14}= 3 \,p_{u \rightarrow v}^2 \, p_{u \leftrightarrow v}$, where
the factor of 3 is the number of possible ways to arrange the pair
connections to obtain the motif (for details and combinatorics of all
patterns see supplementary material \ref{SI-sec:3motif_stat}). The
actual occurrence of motifs in their data, however, differed
significantly from this computed expectation. For example, Song et
al.~found that motif 14 occurred approximately 3 ($\pm 2$) times more
often than expected. Other experimental studies \cite{Perin2011,
  Rieubland2014} and recent computational studies
\cite{Miner2016,Gal2017} have corroborated the specific occurrence of
three neuron patterns in cortical circuits.

Here we analyzed the counts of triplet motifs in anisotropic networks
and compared it with the expected counts predicted from the pair
statistics extracted before. We found a highly specific distribution
of over- and underrepresentation of three neuron motifs in anisotropic
networks (\figref{fig:3_motifs}A) and tuned anisotropic networks
(\figref{fig:3_motifs}B) resembling the experimental results of
\textcite{Song2005}.%
%
\input{figures/figure_4.tex}%
%
Notably, deviation from the expected counts was in general much lower
in rewired and distance-dependent versions of the networks. Indeed,
networks with anisotropy in spatial connectivity have significantly
over- and underrepresented. The biggest 10, 12, 14 and
15. (\figref{fig:3_motifs}, \figref{SI-fig:3motif_rel}).


Perin et al.~reported an overrepresentation of motifs 4, 10, 12 and 14
in connections between layer 5 pyramidal cell in rat somatosensory
cortex. We found that these motifs are strongly overrepresented in
both anisotropic and tuned anisotropic networks (motif labels marked
with *). Notably, deviation from the expected counts is in general
much lower in the rewired and distance-dependent versions of the
networks. Specifically, strong overrepresentation .

Taken together the data show that anisotropy in spatial connectivity
strongly affects the relative occurrence of three neuron connection
patterns and induces an overrepresentation in motifs that were
reported as appearing above expectation in cortical circuits.

% These data match well . Here, over- and underrepresentation of certain
% motifs is a direct result of anisotropy in spatial connectivity as
% rewired or distance-dependent versions of the networks do not show the
% same degree of deviation from the expected motif occurrence.

% The strongest deviations were
% observed in motifs 8,10,12,14,15 and 16 (overrepresented) and motif 9 ??
% (underrepresented). Rewired and distance-dependent networks show a
% deviation from the expectation as well, however, the effect is much
% stronger in anisotropic networks

% Interestingly, this effect is much stronger than in
% distance-dependent, where one might expect significant over- and
% underrepresentations as well. Finding two connected pairs in a triplet
% indicates a spatial closeness, that can be transferred to third pair
% as well - thus one should find a higher connection chance. We find
% that in comparable distance-dependent networks, only



%
% To expose the effect of anisotropy in connectivity on the statistics
% of the occurrence of motifs composed of three neurons, in each of the
% three graph types (anisotropic, rewired, distance-dependent) a sample
% of 3 graphs was taken and in each of those graphs the motifs of 300000
% random triplets of neurons were registered. % song_alt_dist_N3_300k 
% The observed occurrence of three-neuron motifs distinctly differed
% from what one might expect from relative frequencies of connections in
% neuron pairs: Assuming independence, the expected relative frequency
% of the 16 different motifs were calculated from probabilities $p_u,
% p_s$ and $p_r$ from above %?? are they really mentioned abov?.  
% \ref{SI_song_motif_comp}%??
% Displaying the relative counts (counts extracted from graphs divided
% by counts expected from neuron pair statistics) in
% Fig.~\ref{fig_three-neuron}, we find that anisotropy has significant
% impact of distribution of motifs. 

