\clearpage
\pagebreak
\newpage

\subsection*{Neuron clusters}

In the study. The number of connections occurring in clusters
. Densely connected clusters are occurring much more than one would
expect from distance-dependent connectivity.

Here we performed a similar analysis. Does anisotropy in spatial
connectivity have an effect on the number of connections in neuron
clusters? In groups of 3, 6, 8 and 12 randomly selected neurons we
counted the number of connections in anisotropic and tuned anisotropic
networks as well as in their rewired versions. We found that high
counts of connections appear much more frequently in networks with
anisotropy than in rewired networks
(\figref{fig:neuron_clusters}). Anisotropy in spatial connectivity
thus induces an overrepresentation of densely connected neuron groups
as observed experimentally in rat somatosensory cortex.

\input{figures/figure_5.tex}


Notably, this effect does not seem to be sensitive to the particular
way anisotropy in spatial connectivity was implemented as
overrepresentation seemed approximately equal in anisotropic and tuned
anisotropic networks. 


% While rewiring, some connections have been lost (numbers!), that means
% some of the overrepresentation might be due to that as this
% relationship is strict (no new connections get introduced). It was
% shown that small deviations *can* have a significant effenct - see for
% example 85c70a9f\_1v1. However, in average no such is noticeable as
% expected. 

The significant overrepresentation from rewired to distance-dependent
is something important to think about. It kind of destroys the point
I'm trying to make: Anisotropy makes all the things! So, it's very
important to understand: Is it really true that the rewired-distance
effect is strong? (So far this has been only tested for one case:
8counts) Then, it's crucial to understand where this difference is
coming from. What aspect is in rewired, but not in distance, that
produces this?

We tested whether . Not such. Also, while rewiring had a strong effect
on the occurrence of , it did not affect distance-dependent networks
(\figref{SI-fig:nclst})
